### 2012-02-12 exam3/1.tex
add set problem

The Northeastern University Registrar is instituting a new course
registration system, in which each student will wait in a ``Virtual
Line'' until every student ahead of them has registered.  A simple way
to represent a line (also known as a \emph{queue}) is by using a list.
But this representation makes it slow to add somebody to the end of
the line (or to take somebody off the front of the line, depending on
whether the front of the list represents the front or rear of the
line).

In order to provide maximal waiting efficiency, you have been tasked
with implementing a representation that uses \emph{two} lists!  The
key idea of this fancy representation is that one list will represent
some portion of the front of the line, while the other will represent
the remainder of the line \emph{in reverse order}.  So if you're the
first element of the first list, you are at the head of the line.  On
the other hand, if you're the first element of the second list, you
are the very last person in line.
%
Here is the interface for queues:
\begin{verbatim}
;; A [IQ X] implements:
;;
;; head : -> X
;; Produce the element at the head of this queue.
;; Assume: this queue is not empty.
;;
;; deq : -> [IQ X]      (Short for "dequeue")
;; Produces a new queue like this queue, but without 
;; this queue's first element.
;; Assume: this queue is not empty.
;;
;; enq : X -> [IQ X]    (Short for "enqueue")
;; Produce new queue with given element added to the
;; END of this queue.
;;
;; emp? : -> Boolean
;; Is this queue empty?
\end{verbatim}

The \code{head} and \code{deq} operations require that the queue be
non-empty when they are used, but this can be assumed and these
operations do not need to check for an empty queue.

Further, the Registrar's office has just learned about
\emph{invariants}, and insists on maintaining the following invariant
about all of their queues:

\begin{quote}
  {\it if the front of the queue is empty, the whole queue must also be empty.}
\end{quote}

\noindent
The Registrar's office has given you three tasks to prepare their
Virtual Line for its launch later this semester:

\begin{enumerate}
\item Design an implementation of the queue data structure to the Registrar's
  specifications.  You must maintain the invariant stated above, and
  you should take advantage of the invariant when implementing the
  operations.
\item Unfortunately, when testing the queue, the Registrar has
  discovered that some queues with the same elements in the same order
  can be represented in multiple ways.  Give an example of two
  different representations of the same queue.  Implement a
  \code{to-list} operation which produces a list of elements going in
  order from the front to the rear of the queue. In your tests, you
  should show how this addresses the problem.
\item The Registrar has a problem with careless data entry.  Design
  and implement a constructor for queues which, given two input lists
  of elements, ensures that the invariant is maintained.  
\end{enumerate}

\begin{solution}
\begin{verbatim}
;; Part 1:
;; invariant: front is empty IFF queue is empty
;; A [Queue X] is (queue% [Listof X] [Listof X])           [1pt]
;; and implements [IQ X]

(define-class queue%                                       [1pt]
  (fields front back)
  ;; Add element to the back
  ;; X -> [Queue X]
  ;; add it to the front IFF the front is empty
  ;; (otherwise invariant is violated)
  (check-expect (eq . enqueue 1) (queue% '(1) '()))        [1pt]
  (check-expect (q . enqueue 5) (queue% '(1 2 3) '(5)))
  (define/public (enqueue i)
    (cond [(empty? (field front))                          [2pt]
           (queue% (list i) empty)]
          [else
           (queue% (field front) (cons i (field back)))])) [2pt]
  ;; get the first element
  ;;  -> X
  ;; assumes this queue is non-empty
  ;; uses the invariant
  (check-expect (q . head) 1)                              [1pt]
  (define/public (head)                                    
    (first (field front)))
  ;; drop the first element
  ;;  -> [Queue X]
  ;; assumes this queue is non-empty
  (check-expect (eq . enqueue 1 . dequeue) eq)             [1pt]
  (check-expect (q . enqueue 5 . dequeue) (queue% '(2 3) '(5)))
  (define/public (dequeue)
    (cond 
      ;; maintain the invariant -- don't let `front' be empty
      [(empty? (rest (field front)))                       [3pt]
       (queue% (reverse (field back)) empty)]
      ;; otherwise, simple
      [else (queue% (rest (field front)) (field back))]))  [2pt]
  ;; is this queue empty?
  ;; emp? : -> Boolean
  ;; uses the invariant
  (check-expect (q . emp?) false)                          [1pt]
  (check-expect (eq . emp?) true)
  (define/public (emp?)
    (empty? (field front)))

;; Part 2:

;; Examples:                                               [1pt]
(queue% (list 1) (list 6)) vs (queue% (list 1 6) empty)
(queue% (list 1 8 12) (list 6)) vs (queue% (list 1 8) (list 12 6))

  ;; convert the queue to a list                           [1pt]
  ;; -> [Listof X] 
  ;; turn examples above into tests:                       [1pt]
  (check-expect ((queue% (list 1) (list 6)) . to-list)
		((queue% (list 1 6) empty) . to-list))
  (check-expect ((queue% (list 1 8 12) (list 6)) . to-list)
		((queue% (list 1 8) (list 12 6)) . to-list))
  (check-expect (q . to-list) '(1 2 3))                    
  (check-expect (q . enqueue 5 . to-list) '(1 2 3 5))   
  (check-expect (q . enqueue 5 . dequeue . to-list) '(2 3 5))
  (define/public (to-list)                                 [1pt]
    (append (field front)
            (reverse (field back)))))

(define q (queue% '(1 2 3) empty))
(define eq (queue% empty empty))

;; Part 3:                                                 [2pt]
;; several good possibilities (others are possible)
(constructor (f b)
  (fields (append f (reverse b)) empty))

(constructor (f b)
  (cond [(empty? f) (fields (reverse b) empty)]
	[else (fields f b)]))

(constructor (f b)
  (cond [(and (empty? f) (empty? b)) (fields f b)]
	[(empty? f) (fields (reverse b) empty)]
	[else (fields f b)]))

\end{verbatim}
\end{solution}
\end{problem}

@@

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact| method to use them.
You may modify your natural number implementation if neccessary for
your solution.

@@

\ifrubric\else
\newpage
[Here is some more space for the previous problem.]
\fi
\end{problem}

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times n)))

@@

(define true (new t%))
(define false (new f%))
\end{verbatim}
\end{solution}
\end{document}

>>>>

In order to manage class scheduling, the College decided that they
needed to create a new implementation of the \verb|Set| data
structure.  They went with the lowest bidder, which
turned out to be a bunch of BU students.  Now, having paid for this
code, the College needs to test it, and they turn to you.

Unfortunately, code the BU students produced is so convoluted that it
doesn't bear looking at.  All you have to go on is that the \verb|Set|
data structure obeys the following interface:

\begin{verbatim}
;; empty-set is a set

;; A Set implements:
;; - insert : String -> Set
;; adds the given string to the set, if it isn't already there
;; - remove : String -> Set
;; removes the given string from the set
;; if the given string is not in the set, produce the original set
;; - size : -> Number
;; count the elements of the set
;; - has : String -> Boolean
;; is the given String in the set?
\end{verbatim}
\noindent
\verb|Set|s should satisfy the following property, called the
``insert-size'' property, for any element and
set:

\begin{quote}
If we insert an element into a set, then the resulting set is the
same size as the original set \emph{if} the element was already in the
set, and the resulting set is one larger than the original set
if the element was \emph{not} already in the set.
\end{quote}

\begin{enumerate}
\item Codify the ``insert-size'' property as a predicate.  Write a
  test case that should pass (if the sets work right) using the
  predicate you have defined.
\end{enumerate}

\noindent
The College wants to be really sure that their sets are correct, and
they've come up with another property that they think is true, called
the ``insert-remove'' property:

\begin{quote}
If we insert element $a$ into a set, and then remove element $b$ from
the resulting set, the result has the same size as if we do the two
operations in the reverse order.
\end{quote}

\noindent
The College is considering adding this to the test suite as well.

\begin{enumerate}
\setcounter{enumi}{1}

\item Codify this property as a predicate, and write a test case that
  should pass using the predicate.

\item Is the ``insert-remove'' property true for all sets?  If so,
  give an explanation.  If not, write a counter-example (i.e., a
  failing test case) using your predicate.

\begin{solution}
\begin{verbatim}
;; pred1 : Set String -> Boolean
;; check the "insert-size" property
(define (pred1 s i)
  (or (and (s . has i) (= (s . insert i . size) (s . size)))
      (and (not (s . has i))
           (= (s . insert i . size) (add1 (s . size))))))

(check-expect (pred1 empty-set "hi") true)

;; pred2 : Set String -> Boolean
;; check the "remove-insert" property
(define (pred2 s a b)
  (= (s . insert a . remove b . size)
     (s . remove b . insert a . size))

(check-expect (pred2 empty-set "yes" "no"))

;; counter-example
(check-expect (pred2 empty-set "yes" "yes"))
\end{verbatim}
\end{solution}
\end{problem}

@@

Following these interface definitions, design data and class
defintions for implementations of the \verb|Bool| interface, and
reimplement the \verb|fact| method to use them.
You will need to provide new defintions for \verb|true| and
\verb|false|, since they were broken by this catastrophe as well.

@@

\begin{solution}
\begin{verbatim}
(define-class fact%
  (define (fact n)
    (n . zero? . branch (new const% one) (new recur% n))))
(define-class const%
  (fields run))
(define-class recur%
  (fields n)
  (define (run) ((new fact%) . fact (this . n . sub1) . times (this . n))))

@@

(define true (new t%))
(define false (new f%))
\end{verbatim}
\end{solution}
\end{problem}
\end{document}
